Density-dependent selection and the limits of relative fitness

### 2018-02-26 manuscript.tex
l

The approximations \eqref{eq:meanfieldr} and \eqref{eq:meanfielda} must be done in a way that ensures consistency between rare and common types. Suppose that two identical types are present (same $b$, $c$ and $d$), one with low density $l_1\ll 1$, and the other with high density $l_2\approx L\gg 1$. Since $L$ is large, uncontested territories make a negligible contribution each type's growth. The rare type grows almost entirely due to $\Delta_r n_1$, while the common type grows almost entirely due to $\Delta_a n_2$. To ensure consistency, these identical types should retain constant relative frequencies, and so the approximate per-capita growth rates $\Delta_r n_1/m_1 = \Delta_a n_2/m_2$ implied by \eqref{eq:meanfieldr} and \eqref{eq:meanfielda} must be equal.

@@

To be more concrete, suppose that we use $\tilde{p}$ and $\hat{p}$ in place of $\tilde{q}$ and $\hat{q}$ in Eqs.~\eqref{eq:meanfieldr} and \eqref{eq:meanfielda}, respectively. Then, since $\langle x_2\rangle_{\tilde{p}}\approx L$, the right hand side of Eq. \eqref{eq:meanfieldr} is approximately $1/(L+1)$ for the rare type. And, since $l_1\ll 1$ and $L\gg 1$ in Eq.~\eqref{eq:deltr}, we have $\Delta_r n_1 \approx 1/(L+1)$. Similarly, for the common type, $\sum_j \langle x_j\rangle_{\hat{p}}\approx L$ in Eq. \eqref{eq:meanfielda}, and so $\Delta_a n_2 \approx 1/L$. Therefore, the consistency requirement is violated if we naively use the conditional distributions $\tilde{p}$ and $\hat{p}$ derived from the dispersal distribution $p$ to calculate the effective means.

>>>>

The approximations \eqref{eq:meanfieldr} and \eqref{eq:meanfielda} must be done in a way that ensures consistency between rare and common types. Suppose that two identical types are present (same $b$, $c$ and $d$), one with low density $l_1\ll 1$, and the other with high density $l_2\approx L\gg 1$. Since $L$ is large, uncontested territories contribute negligibly to each type's growth. The rare type grows almost entirely due to $\Delta_r n_1$, while the common type grows almost entirely due to $\Delta_a n_2$. To ensure consistency, these identical types should retain constant relative frequencies, and so the approximate per-capita growth rates $\Delta_r n_1/m_1 = \Delta_a n_2/m_2$ implied by \eqref{eq:meanfieldr} and \eqref{eq:meanfielda} must be equal.

@@

For example, suppose that we use $\tilde{p}$ and $\hat{p}$ in place of $\tilde{q}$ and $\hat{q}$ in Eqs.~\eqref{eq:meanfieldr} and \eqref{eq:meanfielda}, respectively. Then, since $\langle x_2\rangle_{\tilde{p}}\approx L$, the right hand side of Eq. \eqref{eq:meanfieldr} is approximately $1/(L+1)$ for the rare type. And, since $l_1\ll 1$ and $L\gg 1$ in Eq.~\eqref{eq:deltr}, we have $\Delta_r n_1 \approx 1/(L+1)$. Similarly, for the common type, $\sum_j \langle x_j\rangle_{\hat{p}}\approx L$ in Eq. \eqref{eq:meanfielda}, and so $\Delta_a n_2 \approx 1/L$. Therefore, the consistency requirement is violated if we naively use the conditional distributions $\tilde{p}$ and $\hat{p}$ derived from the dispersal distribution $p$ to calculate the effective means.
